---
title: "Data manipulation"
author: "Jingqi Song"
date: "September 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

# Manipulation

select:for columns
fill out:for rows

## Import FAS csv files

import my first csv to use as examples

```{r}
litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```

## select variables

```{r}
select(litters_data,litter_number,gd0_weight,pups_survive)

```

...by specifying a range;

```{r}
select(litters_data,gd_of_birth:pups_survive)
select(litters_data,group,gd_of_birth:pups_survive)
```

...by saying what you want to remove;

```{r}
select(litters_data,-litter_number)
```

...rename;

```{r}
select(litters_data,group,litter_number,gest_day_0_weight = gd0_weight)
rename(litters_data,gest_day_0_weight = gd0_weight)
```

...use select helpers

```{r}
select(litters_data,group,starts_with('gd'))
select(litters_data,starts_with('gd'))

select(litters_data,litter_number,everything())
```

learning assessment: pups dataset

```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## filter observations

true or false? if it is ... i'm gonna to keep this thing
==: whether the 2 things are the same

```{r}
filter(litters_data,gd0_weight<25)
filter(litters_data,pups_born_alive == 8)

filter(litters_data,is.na(gd0_weight))
filter(litters_data,!is.na(gd0_weight))

filter(litters_data, group == 'low8' | group == 'low7')
filter(litters_data, group %in% c('low8', 'low7'))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
```

## mutate

create variables
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
)

mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_square = wt_gain^2)
```

## arrange

```{r}
arrange(litters_data,gd0_weight)

arrange(litters_data,desc(pups_born_alive), gd0_weight)
```

## piping

look at intermediate object approach

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars

```

nested function calls

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean

```

use piping


```{r}
read_csv("./data/FAS_litters.csv",  col_types = "ccddiiii")%>%
  janitor::clean_names()%>%
  select(-pups_survive)%>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  )%>%
  filter(!is.na(gd0_weight))
```

